---
title: "What did the presidents say at their inauguation?"
output:
  html_document: default
  html_notebook: default
---

```{r,message=FALSE,include=FALSE}
library(rJava)
library(xlsxjars)
library(xlsx)
library(dplyr)
library(tidytext)
library(NLP)
library(tm)
library(RColorBrewer)
library(wordcloud)
library(beeswarm)
library(tokenizers)
library(ggplot2)
library(plotly)
library(DT)
library(qdap)
library(syuzhet)
library(topicmodels)
```

```{r,message=FALSE,include=FALSE}
#load data
info <- read.xlsx("../data/InaugurationInfo.xlsx",1)
date.file <- read.table("../data/InauguationDates.txt", header=F, sep='\t', fill=T)



#using 'date.file' to generate a 'date.vec1' about time
date.vec1 <- NULL
for(i in 2:nrow(date.file)){
  for(j in 2:ncol(date.file)){
    date.vec1 <- c(date.vec1,as.character(date.file[i,j]))
  }
}
date.vec1 <- date.vec1[!(date.vec1 == "")]
#using 'date.file' to generate a 'date.vec2' about president
date.vec2 <- NULL
for(i in 2:nrow(date.file)){
  n <- sum(date.file[i,2:5] != "")
  date.vec2 <- c(date.vec2,rep(as.character(date.file[i,1]),n))
}
#using 'date.file' to generate a 'date.vec3' about term
date.vec3 <- NULL
for(i in 2:nrow(date.file)){
  n <- sum(date.file[i,2:5] != "")
  date.vec3 <- c(date.vec3,c(1:n))
}
#combine 'date.vec1' and 'date.vec2' and 'date.vec3' into 'date'
date <- cbind(date.vec1,date.vec2,date.vec3)
date <- as.data.frame(date)
names(date) <- c("Time","President","Term")
```

```{r,warning=FALSE,include=FALSE}
setwd("C:/Users/dingx/Documents/GitHub/Spring2018-Project1-XueyingDing/data/InauguralSpeeches")

#generate a 'speech.vec1' about speech
speech.vec1 <- NULL
for(i in 1:58){
  file <- readLines(list.files()[i])
  file <- paste(file,collapse = "")
  speech.vec1 <- c(speech.vec1,file)
}
#generate a 'speech.vec2' about president
#generate a 'speech.vec3' about term
speech.vec2 <- NULL
speech.vec3 <- NULL
for(i in 1:58){
  text.name <- list.files()[i]
  president <- substring(text.name,6,nchar(text.name)-6)
  term <- substring(text.name,nchar(text.name)-4,nchar(text.name)-4)
  term <- as.numeric(term)
  speech.vec2 <- c(speech.vec2,president)
  speech.vec3 <- c(speech.vec3,term)
}
#combine 'speech.vec1' and 'speech.vec2' and 'speech.vec3' into 'speech'
speech <- cbind(speech.vec1,speech.vec2,speech.vec3)
speech <- as.data.frame(speech)
names(speech) <- c("Speech","President","Term")

speech$Term <- as.numeric(speech$Term)
speech$President <- as.character(speech$President)
speech$Speech <- as.character(speech$Speech)
date$Term <- as.numeric(date$Term)
date$President <- as.character(date$President)
info$President <- as.character(info$President)

#combine 'info' and 'date' and 'speech' into a new data frame 'mydata'
mydata <- info %>%
        left_join(date,by = c("President","Term"))
mydata <- merge(mydata,speech,by.x = c("File","Term"),by.y = c("President","Term"))
#sort 'data' by time
mydata$Time <- as.Date(mydata$Time, format="%m/%d/%Y")
mydata <- mydata[order(mydata$Time),]
```
\
\
\
**Have you ever been curious about what do presidents say on their inauguration? Here we have 58 speeches given by 40 USA presidents in history and we will use these data to do a brief study exploring fun facts within it. Let's get started! **
\
\
###What words do they use mostly?
```{r,warning=FALSE,include=FALSE}
#create 'word_cloud' function
word_cloud <- function(article){

docs <- Corpus(VectorSource(article))
  
  
  #text transformation
  toSpace <- content_transformer(function (x , pattern ) gsub(pattern, " ", x))
  docs <- tm_map(docs, toSpace, "/")
  docs <- tm_map(docs, toSpace, "@")
  docs <- tm_map(docs, toSpace, "\\|")
  
  #clean the text
  # Convert the text to lower case
  docs <- tm_map(docs, content_transformer(tolower))
  # Remove numbers
  docs <- tm_map(docs, removeNumbers)
  # Remove english common stopwords
  docs <- tm_map(docs, removeWords, stopwords("english"))
  # Remove your own stop word
  # specify your stopwords as a character vector
  docs <- tm_map(docs, removeWords, c("blabla1", "blabla2"))
  # Remove punctuations
  docs <- tm_map(docs, removePunctuation)
  # Eliminate extra white spaces
  docs <- tm_map(docs, stripWhitespace)
  # Text stemming
  docs <- tm_map(docs, stemDocument)
  
  #Build a term-document matrix
  dtm <- TermDocumentMatrix(docs)
m <- as.matrix(dtm)
v <- sort(rowSums(m),decreasing=TRUE)
d <- data.frame(word = names(v),freq=v)


#Generate the Word cloud
set.seed(1234)
wordcloud(words = d$word, freq = d$freq, min.freq = 1,
          max.words=200, random.order=FALSE, rot.per=0.35,
          colors=brewer.pal(8, "BuGn"))
}
#create 'top10.word' function
top10.word <- function(article){
  
docs <- Corpus(VectorSource(article))

#text transformation
toSpace <- content_transformer(function (x , pattern ) gsub(pattern, " ", x))
docs <- tm_map(docs, toSpace, "/")
docs <- tm_map(docs, toSpace, "@")
docs <- tm_map(docs, toSpace, "\\|")

#clean the text
# Convert the text to lower case
docs <- tm_map(docs, content_transformer(tolower))
# Remove numbers
docs <- tm_map(docs, removeNumbers)
# Remove english common stopwords
docs <- tm_map(docs, removeWords, stopwords("english"))
# Remove your own stop word
# specify your stopwords as a character vector
docs <- tm_map(docs, removeWords, c("blabla1", "blabla2"))
# Remove punctuations
docs <- tm_map(docs, removePunctuation)
# Eliminate extra white spaces
docs <- tm_map(docs, stripWhitespace)
# Text stemming
#docs <- tm_map(docs, stemDocument)

#Build a term-document matrix
dtm <- TermDocumentMatrix(docs)
m <- as.matrix(dtm)
v <- sort(rowSums(m),decreasing=TRUE)
####d <- data.frame(word = names(v),freq=v)
d <- data.frame(word = names(v))
return(head(d,10))
}
```
\
\
**Here we compare presidents into two most famous parties -- Democratic and Republican.**
\
```{r,echo=FALSE,fig.height=4,fig.align='center'}
par(mfrow=c(1,2),mar = c(3, 0, 2, 0))
party <- unique(mydata$Party)

  mytext <- paste(mydata$Speech[mydata$Party == party[4]],collapse = "")
  word_cloud(mytext)
  title(main = as.character(party[4]))

  mytext <- paste(mydata$Speech[mydata$Party == party[6]],collapse = "")
  word_cloud(mytext)
  title(main = as.character(party[6]))
```
\
**Hmm... It seems that there's no big difference between most frequently used words by these two parties. They all focus on 'people', 'govern', 'nation', etc. Since these are inauguation speeches, presidents would like to say many 'will' to illustrate what they gonna do in their terms of office.**
\
**In order to better understanding, you can check top 10 words used in every year here.**
####Democratic
```{r,echo=FALSE}
#generate data frame 'Democratic.df'
Time_D <- as.data.frame(mydata$Time[mydata$Party == "Democratic"])
Speech_D <- as.data.frame(mydata$Speech[mydata$Party == "Democratic"])
Word_D <- apply(Speech_D,1,top10.word)
Word_D <- as.data.frame(Word_D)
Word_D <- t(Word_D)
Democratic.df <- cbind.data.frame(Time_D,Word_D)
names(Democratic.df) <- c("Year",1:10)
rownames(Democratic.df) <- NULL

#generate data frame 'Republican.df'
Time_R <- as.data.frame(mydata$Time[mydata$Party == "Republican"])
Speech_R <- as.data.frame(mydata$Speech[mydata$Party == "Republican"])
Word_R <- apply(Speech_R,1,top10.word)
Word_R <- as.data.frame(Word_R)
Word_R <- t(Word_R)
Republican.df <- cbind.data.frame(Time_R,Word_R)
names(Republican.df) <- c("Year",1:10)
rownames(Republican.df) <- NULL
datatable(Democratic.df, options = list(pageLength = 5))
```
####Republican
```{r,echo=FALSE}
datatable(Republican.df, options = list(pageLength = 5))
```
\
\
\
##What are presidents' speech-style?
\
\
```{r,warning=FALSE,include=FALSE}
#generating 'sentence.list2'
sentence.list2=NULL
for(i in 1:nrow(mydata)){
  sentences=sent_detect(mydata$Speech[i],
                        endmarks = c("?", ".", "!", "|",";"))
  if(length(sentences)>0){
    emotions=get_nrc_sentiment(sentences)
    word.count=word_count(sentences)
    # colnames(emotions)=paste0("emo.", colnames(emotions))
    # in case the word counts are zeros?
    emotions=diag(1/(word.count+0.01))%*%as.matrix(emotions)
    sentence.list2 <- rbind(sentence.list2, 
                        cbind(mydata[i,-ncol(mydata)],
                              sentences=as.character(sentences), 
                              word.count,
                              emotions,
                              sent.id=1:length(sentences)
                              )
    )
  }
}

sentence.list2=
  sentence.list2%>%
  filter(!is.na(word.count))

#generating 'sentence.list3'
sentence.list3 <- data.frame()
P.T <- paste(sentence.list2$President,sentence.list2$Term,sep = "")
P.T <- as.data.frame(P.T)
sentence.list3<- cbind.data.frame(P.T,sentence.list2$President,
                                  sentence.list2$word.count,
                                  as.numeric(as.character(sentence.list2$Words)))
names(sentence.list3)[2:4] <- c("President","word.count","Words")
```
**One of most direct way to inspect presidents' speech style is figuring out how many words do they use in speech,**
\
\
```{r,message=FALSE,echo=FALSE,fig.align='center'}
p <- ggplot(data = sentence.list3, aes(x = President, y = Words, fill = P.T)) +
              geom_col(position = "dodge") + 
              theme(axis.text.x = element_blank(),
                    axis.ticks.x = element_blank())
ggplotly(p)
```
\
**Move your mouse, you can check whatever you want! We notice that some presidents give speeches of very different length in their several terms. Do they change their writing style during terms? Let's do more research on how they use words.**
**First, we want to inspect President Franklin D. Roosevelt. He is the only one who has 4 terms!**
\
```{r,message=FALSE,echo=FALSE,fig.align='center'}
#Franklin D. Roosevelt
FranklinDRoosevelt.df <- sentence.list3[sentence.list3$President 
                                        == "Franklin D. Roosevelt",c(1:3)]
FranklinDRoosevelt.df$President <- as.character(FranklinDRoosevelt.df$President)
FranklinDRoosevelt.df$P.T <- as.factor(as.character(FranklinDRoosevelt.df$P.T))
#Franklin D. Roosevelt's beeswarm plot
beeswarm(word.count ~ President, data = FranklinDRoosevelt.df,
    method = 'swarm',
    pch = 16, pwcol = as.numeric(P.T) + 4)
legend('topright', legend = levels(FranklinDRoosevelt.df$P.T),
           title = 'President Term', pch = 16, col = 5:8,cex = 0.75)
```
**In this plot, different colors mean different term and y-axis means how many words are used in each sentence. We can know that his style doesn't change so much because dots in every color have similar distribution according to y-axis.**
**How about these guys? Abraham Lincoln, George Washington, James Monroe, William McKinley.**
```{r,echo=FALSE,fig.height=4,fig.align='center'}
#generating data frame about presidents(1) -- Abraham Lincoln
AbrahamLincoln.df <- sentence.list3[sentence.list3$President 
                                        == "Abraham Lincoln",c(1:3)]
AbrahamLincoln.df$President <- as.character(AbrahamLincoln.df$President)
AbrahamLincoln.df$P.T <- as.factor(as.character(AbrahamLincoln.df$P.T))
#generating data frame about presidents(2) -- George Washington
GeorgeWashington.df <- sentence.list3[sentence.list3$President 
                                        == "George Washington",c(1:3)]
GeorgeWashington.df$President <- as.character(GeorgeWashington.df$President)
GeorgeWashington.df$P.T <- as.factor(as.character(GeorgeWashington.df$P.T))
#generating data frame about presidents(3) -- James Monroe
JamesMonroe.df <- sentence.list3[sentence.list3$President 
                                        == "James Monroe",c(1:3)]
JamesMonroe.df$President <- as.character(JamesMonroe.df$President)
JamesMonroe.df$P.T <- as.factor(as.character(JamesMonroe.df$P.T))
#generating data frame about presidents(4) -- William McKinley
WilliamMcKinley.df <- sentence.list3[sentence.list3$President 
                                        == "William McKinley",c(1:3)]
WilliamMcKinley.df$President <- as.character(WilliamMcKinley.df$President)
WilliamMcKinley.df$P.T <- as.factor(as.character(WilliamMcKinley.df$P.T))

#plot
par(mfrow = c(1,2))
#plot(1) -- Abraham Lincoln
beeswarm(word.count ~ President, data = AbrahamLincoln.df,
    method = 'swarm',
    pch = 16, pwcol = as.numeric(P.T) + 4,xlab = '')
legend('topright', legend = levels(AbrahamLincoln.df$P.T),
           title = 'President Term', pch = 16, col = 5:6,cex = 0.45)
#plot(2) -- George Washington
beeswarm(word.count ~ President, data = GeorgeWashington.df,
    method = 'swarm',
    pch = 16, pwcol = as.numeric(P.T) + 4,xlab = '')
legend('topright', legend = levels(GeorgeWashington.df$P.T),
           title = 'President Term', pch = 16, col = 5:6,cex = 0.45)
par(mfrow = c(1,2))
#plot(3) -- James Monroe
beeswarm(word.count ~ President, data = JamesMonroe.df,
    method = 'swarm',
    pch = 16, pwcol = as.numeric(P.T) + 4,xlab = '')
legend('topright', legend = levels(JamesMonroe.df$P.T),
           title = 'President Term', pch = 16, col = 5:6,cex = 0.45)
#plot(4) -- William McKinley
beeswarm(word.count ~ President, data = WilliamMcKinley.df,
    method = 'swarm',
    pch = 16, pwcol = as.numeric(P.T) + 4,xlab = '')
legend('topright', legend = levels(WilliamMcKinley.df$P.T),
           title = 'President Term', pch = 16, col = 5:6,cex = 0.45)
```
\
**It's clear that Abraham Lincoln's speech length change much and George Washington's words using style cahnge much.**
\
**We are also attracted to highest bar in former bar chart -- William Henry Harrison. Why his speech has so many words? We'll see.**
```{r,echo=FALSE}
j2 <- sentence.list2%>%
  filter(File=="WilliamHenryHarrison", 
         word.count>=80)%>%
  select(sentences)
datatable(j2, options = list(pageLength = 5))
```
\
**Amazing! Usually speaking, centence consist of 10-15 words but we can see that he even has many sentences consist of over 80 words!**
\
**We also notice that there's vacancy in former bar chart. After checking data, we can tell you that's given by Donald J. Trump. He seems like using short expression so let's look at his sentences consist of under 10 words.**
```{r,echo=FALSE}
j3 <- sentence.list2%>%
  filter(File=="DonaldJTrump", 
         word.count<=10)%>%
  select(sentences)
datatable(j3, options = list(pageLength = 5))
```
\
**It's much easier to read than former table, right?**
\
\
\
##Sentiment Analysis based on Speech
\
\
**Tell you a truth, I like movie 'Inside Out' very much. How about we analysis all speeches based on same color theme -- yellow for joy, purple for fear, red for anger, blue for sadness, green for disgust.**
**By the way, highly recommend this movie :D**
```{r,echo=FALSE,fig.align='center'}
emo.means=colMeans(select(sentence.list2, anger:trust)>0.01)
col.use=c("red2", "darkgoldenrod1", 
            "chartreuse3", "blueviolet",
            "darkgoldenrod2", "dodgerblue3", 
            "darkgoldenrod1", "darkgoldenrod1")
barplot(emo.means[order(emo.means)], las=2, col=col.use[order(emo.means)], horiz=T, main="Speeches")
```
\
**Of course, joy dominates the whole. Unfortunately, things not always go well. For example, Great Depression happens just after Coolidge Prosperity. How speeches are change with history?**
\
\
```{r,echo=FALSE,fig.align='center'}
#write a function 'f.plotsent.len2'
f.plotsent.len2 <- function(In.list, InFile, InTerm, President){
  
  col.use=c("lightgray", "red2", "darkgoldenrod1", 
            "chartreuse3", "blueviolet",
            "darkgoldenrod2", "dodgerblue3", 
            "darkgoldenrod1", "darkgoldenrod1",
            "black", "darkgoldenrod2")
  
  In.list$topemotion=apply(select(In.list, 
                                        anger:positive), 
                                 1, which.max)
  In.list$topemotion.v=apply(select(In.list,
                                          anger:positive), 
                                   1, max)
  In.list$topemotion[In.list$topemotion.v<0.05]=0
  In.list$topemotion=In.list$topemotion+1
  
  temp=In.list$topemotion.v
  In.list$topemotion.v[temp<0.05]=1
  
  df=In.list%>%filter(File==InFile, Term==InTerm)%>%
    select(sent.id, word.count, 
           topemotion, topemotion.v)
  
  ptcol.use=alpha(col.use[df$topemotion], sqrt(sqrt(df$topemotion.v)))
  
  plot(df$sent.id, df$word.count, 
       col=ptcol.use,
       type="h", #ylim=c(-10, max(In.list$word.count)),
       main=President)
}

#using 'f.plotsent.len2' function
#before and after the great depression
par(mfrow = c(2,1))
f.plotsent.len2(In.list = sentence.list2, InFile = "HerbertHoover", 
                InTerm = 1, President = "Herbert Hoover")
f.plotsent.len2(In.list = sentence.list2, InFile = "FranklinDRoosevelt", 
                InTerm = 1, President = "Franklin D. Roosevelt")
```
\
**Herbert Hoover becomes president at the end of Coolidge Prosperity while Franklin D. Roosevelt becomes president at the biginning of Great Depression. We can see Herbert Hoover's speech has more positive sentiment.**
**What does Franklin D. Roosevelt say in his first term inauguation speech?**
\
\
```{r,echo=FALSE}
speech.df=tbl_df(sentence.list2)%>%
  filter(File=="FranklinDRoosevelt", Term ==1, word.count>=5)%>%
  select(sentences, c(anger:positive))
speech.df=as.data.frame(speech.df)
FranklinDRoosevelt_1 <- as.character(speech.df$sentences[apply(speech.df[,-1], 2, which.max)])
FranklinDRoosevelt_1 <- as.data.frame(FranklinDRoosevelt_1)
datatable(FranklinDRoosevelt_1, options = list(pageLength = 5))
```
\
**He does talk about Great Expression! It's so cool to combine data with vivid history events!**
\